A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae

### 2021-03-17 thesis_chapter_3.Rmd
chapter 3 files updated

---
title: 'A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalage are one of the most important natural dispersal mechanisms in the ocean, and are regarded as the tumble-weed of the ocean. Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown. Past studies also tend to focus on kelp-rafts which can vary greatly in size and making it difficult to form a concensus as to the role of wind and surface currents. In addition past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam. The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectory of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, as well as the role of drag, in macrolagae trajectory and accumulation. This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scnearios causes accumulation in eddies and vortices within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies and vortices, as important surface accumlation zones for solitary floating macroalage."
---

@@

```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2:Diagram of the *E.maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}

@@

p_RK4_July$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

p_RK4_Oct$type <- "RK4" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("32","33","34","35","36","37","338","39","40","41","42","43"), common.legend = TRUE, legend = c("none"))

@@

```
## Comparison of distances

```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}
## July

pos_July <- c("RK4_July",
         "Min_Kfloat_H100_W00_July",
         "Min_Kfloat_H095_W05_July",
         "Min_Kfloat_H090_W10_July",
         "Min_Kfloat_H085_W15_July",
         "Mean_Kfloat_H100_W00_July",
         "Mean_Kfloat_H095_W05_July",
         "Mean_Kfloat_H090_W10_July",
         "Mean_Kfloat_H085_W15_July",
         "Max_Kfloat_H100_W00_July",
         "Max_Kfloat_H095_W05_July",
         "Max_Kfloat_H090_W10_July",
         "Max_Kfloat_H085_W15_July")

@@

custom_labs_July <- c("32","33","34","35","36","37","338","39","40","41","42","43") #create custom labels and add them to dis_plot

@@

## Oct

pos_Oct <- c("RK4_Oct",
         "Min_Kfloat_H100_W00_Oct",
         "Min_Kfloat_H095_W05_Oct",
         "Min_Kfloat_H090_W10_Oct",
         "Min_Kfloat_H085_W15_Oct",
         "Mean_Kfloat_H100_W00_Oct",
         "Mean_Kfloat_H095_W05_Oct",
         "Mean_Kfloat_H090_W10_Oct",
         "Mean_Kfloat_H085_W15_Oct",
         "Max_Kfloat_H100_W00_Oct",
         "Max_Kfloat_H095_W05_Oct",
         "Max_Kfloat_H090_W10_Oct",
         "Max_Kfloat_H085_W15_Oct")

@@

custom_labs_Oct <- c("44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot

@@

```{r distance plot additional, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 10: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring). The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}

>>>>

---
title: 'A numerical experiment investigating the influence of drag in trajectory patterns of floating macroalgae'
author: 
- R Coppin^[University of the Western Cape], C Rautenbach^[National Institute of Water and Atmospheric Research], AJ Smit^[University of the Western Cape]
output:
  word_document:
    reference_docx: markdown/frontiers_template.docx
  html_document:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
bibliography: references.bib
csl: markdown/frontiers.csl.txt
abstract: "**Abstract**: The ocean currents act as an important dispersal mechanism of material on the ocean surface, both natural and anthropogenic. Macroalage are one of the most important natural dispersal mechanisms, and are regarded as the tumble-weed of the ocean. Despite the many studies that exist, the relative role of wind and surface currents in macrolage trajectory is still unknown. Past studies also focus on kelp-rafts, which can vary greatly in size making it difficult to form a concensus as to the relative role of wind versus surface currents. In addition, past studies have not conisidered surface-area in relation to drag characterisics of floating macroalgae, which have been shown to play a role in the trajectory and accumulation of other flotsam. The advances in Lagrangian trajectory modeling has been exponential in recent years and allows the use of numerical experiments to investigate trajectory of flotsam in the ocean. This study aimed to shed light on the relative role of wind versus surface currents, and the role of drag in macrolagae trajectory and accumulation. This study did not investigate kelp-raft behaviour but instead focused on solitary kelp individuals. This was achieved by comparing simulations of virtual kelp particles which incorporate drag with that of purley Lagrangian particles. The results showed that the virtual kelp particles do follow the ocean currents in general, but that the inclusion of drag into the scenarios causes accumulation in eddies within the study domain. The results also point to the importance of sub-mesoscale features, such as eddies, as important surface accumlation zones for solitary floating macroalage."
---

@@

```{r message=FALSE, warning=FALSE, include=TRUE, echo=FALSE, fig.height=4, fig.width=7, fig.cap="Figure 2: Diagram of the *E.maxima* morphological characteristics considered in this study, along with the associated geometric shapes used to calculate the overall surface area of each morphological characteristic."}

@@

p_RK4_July$type <- "RK4_July" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

p_RK4_Oct$type <- "RK4_Oct" # add extra column 'type' to dataframe and assigned 'RK4' label

@@

figure_a_July <- ggarrange(heatmap_Min_Kfloat_H100_W00_July, heatmap_Min_Kfloat_H095_W05_July,heatmap_Min_Kfloat_H090_W10_July,heatmap_Min_Kfloat_H085_W15_July, heatmap_Mean_Kfloat_H100_W00_July, heatmap_Mean_KFloat_H095_W05_July, heatmap_Mean_KFloat_H090_W10_July, heatmap_Mean_KFloat_H085_W15_July, heatmap_Max_KFloat_H100_W00_July,heatmap_Max_KFloat_H095_W05_July, heatmap_Max_KFloat_H090_W10_July, heatmap_Max_KFloat_H085_W15_July, labels = c("31","32","33","34","35","36","37","38","39","40","41","42"), common.legend = TRUE, legend = c("none"))

@@

```{r distance plot processing additional, message=FALSE, warning=FALSE, include=FALSE}

pos_July <- c("RK4_July",             ## July
         "Min_Kfloat_H100_W00_July",
         "Min_Kfloat_H095_W05_July",
         "Min_Kfloat_H090_W10_July",
         "Min_Kfloat_H085_W15_July",
         "Mean_Kfloat_H100_W00_July",
         "Mean_Kfloat_H095_W05_July",
         "Mean_Kfloat_H090_W10_July",
         "Mean_Kfloat_H085_W15_July",
         "Max_Kfloat_H100_W00_July",
         "Max_Kfloat_H095_W05_July",
         "Max_Kfloat_H090_W10_July",
         "Max_Kfloat_H085_W15_July")

@@

custom_labs_July <- c("30","31","32","33","34","35","36","37","38","39","40","41", "42") #create custom labels and add them to dis_plot

@@

pos_Oct <- c("RK4_Oct",              ## Oct
         "Min_Kfloat_H100_W00_Oct",
         "Min_Kfloat_H095_W05_Oct",
         "Min_Kfloat_H090_W10_Oct",
         "Min_Kfloat_H085_W15_Oct",
         "Mean_Kfloat_H100_W00_Oct",
         "Mean_Kfloat_H095_W05_Oct",
         "Mean_Kfloat_H090_W10_Oct",
         "Mean_Kfloat_H085_W15_Oct",
         "Max_Kfloat_H100_W00_Oct",
         "Max_Kfloat_H095_W05_Oct",
         "Max_Kfloat_H090_W10_Oct",
         "Max_Kfloat_H085_W15_Oct")

@@

custom_labs_Oct <- c("43","44","45","46","47","48","49","50","51","52","53","54","55") #create custom labels and add them to dis_plot

@@

```{r distance plot additional, include = FALSE, message=FALSE, warning=FALSE, echo=FALSE, fig.cap="Figure 10: Boxplots of total distance travelled by all particles from the release site over the course of the simulation for each scenario for July (winter) and October (spring). The x-axis represents the scenario number and the y-axis is the distance travelled in kilometers. The lower and upper hinges correspond to the first and third quartiles. The whiskers represent the range, solid black lines represent the median and black dots are outliers. The reader is referred to tables 1 and 2 for details pertaining to each scenario."}
